---
title: "Assignment 3 Task 3"
author: "Bri Baker"
date: "October 30, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```


# Load Data and Clean

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% # read in data
  clean_names() %>% # lower snake case column names
  rename(size_cm = size) # rename size colname to make more descriptive
```


# Section 1. Size comparisons, purple and red urchins

## Exploratory data visualization

```{r}
ggplot(urchins, aes(size_cm)) + # plot urchins by size
  geom_histogram() + # use a histogram
  facet_wrap(~common_name) + # split by common name
  labs(x = "Size (cm)", # lable x axis
       y = "Individuals (count)", #label y axis
       title = "Figure 1: Exploratory histogram of urchin size") # title
```


```{r}
ggplot(urchins, aes(common_name, size_cm)) + #plot urchins by size and name
  geom_boxplot() + # use box plot
  labs(x = "", # no x lab
       y = "Size (cm)", # label y axis
       title = "Figure 2: Exploratory box plot of urchin size")# title

```


## Descriptive statistics table

```{r}
summary_urchins <- urchins %>% # start with urchins
  group_by(common_name) %>%  # recognize common name as category
  summarize(mean_cm = mean(size_cm), # summarize summary stats by common name
            median_cm = median(size_cm), 
            sd_cm = sd(size_cm), 
            count = n())

kable(summary_urchins, "html") %>% #make a table
  kable_styling("striped") #use striped styling
  
```


## Summary Explanation

Figure 1 indicates that there is a fairly normal distribution in size of both species; however, the breaks in the distributions indicate that there may be more complexity at play. Figure 2 demonstrates that red urchins are generally larger than purple urchins, but that the spread of sizes within the red urchin population is larger with a difference in mean of 2 cm (41% difference). That is also indicated by the tighter distribution of purple urchins in Figure 1 that peaks further to the left of the red urchin distri. 


# Section 2. Purple urchin counts by site and year

## Make subset and summarize counts
```{r}
purple_urchins_annually <- urchins %>% # start with urchins
  filter(common_name == "Purple Urchin") %>% # subset purple urchins
  mutate(date = mdy(date), # format date and make "date" class
         year = year(date), # isolate year from date in new column
         site_full = case_when( # translate site abbriviations to full names in new column
           site == "NAPL" ~ "Naples Reef", 
           site == "AQUE" ~ "Arroyo Quemado Reef",
           site == "CARP" ~ "Carpinteria Reef", 
           site == "MOHK" ~ "Mohawk Reef", 
           site == "IVEE" ~ "Isla Vista Reef"
         )) %>% 
  count(site_full , year) # count urchins per site per year
```


##Exploratory graph of purple urchin counts over time

```{r}
ggplot(purple_urchins_annually, aes(year, n)) + # make a graph of purple urchins count/year
  geom_line() + # use a line graph
  facet_wrap(vars(site_full)) + #separate by site
  scale_x_continuous(breaks = seq(2008, 2020, 2)) + # make breaks even
  theme(axis.text.x = element_text(angle = 45)) + # angle labs
  labs(x = "Year", # label and title
       y = "Individuals", 
       title = "Figure 3: Exploratory graph of annual purple urchin counts")
```

