---
title: "Task 4"
author: "Bri Baker"
date: "November 1, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```


# Set up and visualize

## Read in and subset urchins data

I will be looking specifically at purple urchins in  Naples Reef in January 2020, so I will need to read in the urchins dataset and subset it accordingly. Additionally, I will clean attribute names and add date columns to make comparison easier.

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% # read in data
  clean_names() %>% # lower snake case column names
  rename(size_cm = size)  # rename size colname to make more descriptive
  
naples_purple_10 <- urchins %>% # start with urchins
  mutate(date = mdy(date), # format date and make "date" class
         month = month(date), # isolate month from date in new column
         year = year(date), # isolate year from date in new column
         site_full = case_when( # translate site abbriviations to full names in new column
           site == "NAPL" ~ "Naples Reef", 
           site == "AQUE" ~ "Arroyo Quemado Reef",
           site == "CARP" ~ "Carpinteria Reef", 
           site == "MOHK" ~ "Mohawk Reef", 
           site == "IVEE" ~ "Isla Vista Reef"),
         treatment = case_when(
           treatment == "ANNUAL" ~ "Annual",
           treatment == "CONTROL" ~ "Control"
         )) %>%  
  filter(common_name == "Purple Urchin", # subset purple urchins
          site == "NAPL", # from Naples Reef
          year == 2010 , # in 2020
         month == 1)# %>% #in January
  
```


# Visualize

I will utilize histograms and QQ-plots to understand the spread of the data and to assess normality.

```{r}
ggplot(naples_purple_10, aes(size_cm)) + # plot urchins by size
  geom_histogram(bins = 10) + # use a histogram
  facet_wrap(~treatment) + # split by common name
  labs(x = "Size (cm)", # lable x axis
       y = "Individuals (count)", #label y axis
       title = "Figure 1: Exploratory histogram of urchin size") # title
```


```{r}
ggplot(naples_purple_10, aes(sample = size_cm)) + # plot urchins by size
  geom_qq() + # use a histogram
  facet_wrap(~treatment) + # split by common name
  labs(x = "Theoretical Size (cm)", # lable x axis
       y = "Actual Size (cm)", #label y axis
       title = "Figure 2: Exploratory QQ-Plot of urchin size") # title
```


The histograms in Figure 1 indicate that the sizes measured in both treatments likely have normal distributions that are positilvely skewed. It is difficult to say for certain due to the discrete nature of the measurements (taken in 0.5 cm incriments). The positive, roughly linear trend in the QQ plots of Figure 2 helps to clarify normality, with the tails on both ends signalling the skew in these data.


## Summary Statistics 

```{r}
summary_naples_purple_10 <- naples_purple_10 %>% # start with urchins
  group_by(treatment) %>%  # recognize common name as category
  summarize(mean_cm = mean(size_cm), # summarize summary stats by common name
            median_cm = median(size_cm), 
            sd_cm = sd(size_cm), 
            count = n())

kable(summary_naples_purple_10, "html" , caption = "Table 1: Descriptive statistics") %>% #table
  kable_styling("striped") #use striped styling
```


# t-test

```{r}
annual_naples <- naples_purple_10 %>%  # subset naples purple
  filter(treatment == "Annual") %>% # by "Annual"
  pull(size_cm) # make a vector

control_naples <- naples_purple_10 %>% #subset naples purple
  filter(treatment == "Control") %>% # by "Control"
  pull(size_cm) # make a vector


t.test(x = annual_naples, y = control_naples) # run two sample t-test

```


# Explanation

The annual mean of 3.9 cm and control mean of 3.79 cm are not very different from one another. They are 0.19 cm different which translates to a 4.99% difference. However, the medians of the groups (annual = 4 cm, control = 3.5 cm) are 0.5 cm different with a 13% difference and the standard deviations (annual = 0.97 cm, control = 0.87 cm) also indicate that the spread of the control group is tighter than that of the annual group. These statistics indicate that the groups are not exactly the same. Based on the p-value (p = 0.25), the liklihood that samples at least this different as were drawn from the same population is 25%. 
